Solar wind predictions for the Parker Solar Probe orbit

### 2017-10-09 paperMVVB_referee/referee_report_responses.txt

- adopted
+ highlighted
+ can be adopted

Introduction, line 8: “Until today” is unnecessary – suggest removing. SOHO is also at L1 and returns solar wind observations from CELIAS. But maybe it’s not necessary to list the spacecraft, especially as other spacecraft not listed contributed earlier observations to OMNI.

- removed
+ highlighted
+ can be removed

- SOHO added
+ highlighted
+ add SOHO?
+ SOHO/CELIAS not real-time (ca. +3 hours)
+ SOHO/CELIAS does not contribute to OMNI data set

@@

- adopted
+ highlighted
+ can be adopted

- changed 'lag time' to 'time lag'
+ below, change 'lag time' to 'time lag'

Page 10, Table 2: Similar comments to those for Table 1 apply.

- table caption equation references added
- in section 3 sentence with equation (9) clarified

@@

Below equation 8: I don't understand what "with the implementation of these relations into the lognormal function" means exactly. How are the fit coefficients obtained?

+ highlighted
+ explain

Page 13, Table 3: Give the equation for the fit, to place the parameters in context.

- referenced the fit equation
+ highlighted
+ link the fit equation

- referenced fit equations in all other table captions
+ reference fit equations for all other tables

@@

Page 16, line below figure: Not sure what "established" means here. Remove?

- removed
+ highlighted
+ remove

Line 3: What parameter is used for the power law fits? Presumably it's the radial distance, but this could be stated specifically.

- stated specifically and added equation reference
+ highlighted
+ state specifically

@@

Page 17, figure caption: solar wind parameter power exponents for the dependence on radial distance

- adopted
+ highlighted
+ can be adopted

Last sentence of paragraph, suggest: Since this separation is highly variable, and requires significant effort to calculate for an extended time series, we have ignored this aspect in this analysis.

- adopted
+ highlighted
+ adopt reworded

- a few lines before: added 'and' between references 'Bruno et al. (1986) and Balogh et al. (1999)'

### 2017-10-09 paperMVVB/sections/content.tex
adopted referee comments; made changes bold face

\section{Introduction}
%solar wind beginnings
From observations of cometary tail fluctuations \citet{Biermann1951} inferred the presence of a continuous flow of particles from the Sun. With his theoretical solar wind model \citet{Parker1958} formulated the existence of the solar wind even before the first satellites measured it in situ in 1959 \citep{Gringauz1960,Neugebauer1966}.
%in situ spacecraft
The idea of a space mission flying through the solar corona dates back to the founding year of NASA in 1958 \citep{McComas2008}. Since then several space missions have measured the solar wind in situ at a wide range of heliocentric distances, in case of Voyager~1 as far away as \SI{138}{\au}\footnote{\url{https://voyager.jpl.nasa.gov/}} (July 2017), having even left the heliospause into interstellar space at a distance of \SI{121}{\au} \citep{Gurnett2013}.
\textbf{Various} spacecraft have provided a wealth of solar wind measurements near Earth’s orbit, with WIND \citep{Lepping1995,Ogilvie1995}, ACE \citep{Stone1998} and DSCOVR \citep{Burt2012} \hl{still orbiting } around the L1 point \SI{1.5}{million\,\km} ahead of Earth in the sunward direction. Additional measurements at other solar distances were provided by planetary missions to Venus and Mercury, such as PVO \citep{Colin1980} or MESSENGER \citep{Belcher1991}. Ulysses was the first probe that orbited the Sun out of the ecliptic plane and thus could measure solar wind even at polar latitudes \citep{McComas1998}. \textbf{The in situ solar wind measurements closest to the Sun to date were made by} the Helios mission. \textbf{Helios~1, launched in 1974, reached distances of \SI{0.31}{\au}. Helios~2, launched two years later, approached the Sun as close as} \SI{0.29}{\au} \citep{Rosenbauer1977}.
%Solar Probe Plus mission
The NASA Parker~Solar~Probe\footnote{\url{http://parkersolarprobe.jhuapl.edu/}} (PSP), formerly Solar~Probe~Plus, with a planned launch date in mid 2018, will reach after six years in 2024 its closest perihelia at a distance of 9.86~solar radii (\Rs), that is, \SI{0.0459}{\au} \citep{Fox2015}. This distance will be achieved through seven Venus gravity assists with orbital periods of 88--168~days. In its prime mission time 2018--2025 PSP provides 24~orbits with perihelia inside \SI{0.25}{\au} \citep{Fox2015}. Even its first perihelion, 93~days after launch in 2018, will take PSP to an unprecedented distance of \SI{0.16}{\au} (\SI{35.7}{\Rs}). In comparison, the ESA Solar Orbiter mission with a planned launch in February 2019 will have its closest perihelia at \SI{0.28}{\au} \citep{Muller2013}.

@@

Fig.~\ref{fig:OMNI_yearly_ssn_correlation_c_plot} shows yearly medians of the solar wind parameters and the yearly SSN together with the solar cycle number. The reason for correlating the SSN to the solar wind median values is because the position of a lognormal function is defined by its median. The data are averaged to yearly values to avoid seasonal effects during the Earth’s orbit around the Sun caused by its variations in solar latitude and distance.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/OMNI_yearly_ssn_correlation_c_plot.pdf}}
	\caption{The solar wind parameter yearly medians derived from OMNI data and the yearly SSN from the \citet{sidc} with solar cycle number (top). Their correlation coefficients with the yearly SSN are calculated for time lags back to -15 years (bottom).}
	\label{fig:OMNI_yearly_ssn_correlation_c_plot}
\end{figure}

The solar wind velocity, density and temperature depend on the state of the solar cycle \citep{Schwenn1983}. %density modulation with SSN \citep{Schwenn1983} p.~499\\
For instance the fast solar wind \textbf{occurs at times when polar coronal holes extend to lower latitudes, a typical feature of the declining phase of the solar cycle} as pointed out by \citet[p.~75, Figure~3.52]{Bothmer2007}. Therefore the solar wind velocity, density and temperature maxima exhibit \textbf{time lags relative} to the SSN maxima.

The correlation coefficients of the solar wind parameters with the yearly SSN shown in the bottom part of Fig.~\ref{fig:OMNI_yearly_ssn_correlation_c_plot} are calculated for time lags back to \num{-15}~years to cover a time span longer than a solar cycle. As expected, \hl{the correlation coefficients' amplitudes of all parameters decline with increasing time lag } and show a \textbf{period} of about 11~years. The highest correlation coefficient of 0.728 to the SSN is found for the magnetic field strength, it has no time lag. This finding is anticipated because the SSN is found to be directly proportional to the evolution of the photospheric magnetic flux \citep{Smith2003}.
Velocity and temperature show a \textbf{time lag} of 3~years with peak correlation coefficients of 0.453 and 0.540. The density with a correlation coefficient of 0.468 has a time lag of 6~years, which is in agreement with the by \citet{Bougeret1984} reported density anticorrelation to the SSN.	%p.~406\\
% maximal correlation coefficients and their time lags:\\
% magnetic field strength: lag 0, 0.728\\
% velocity: lag -3, 0.453\\
% density: lag -6, 0.468\\
% temperature: lag -3, 0.540\\

%\subsection{SSN fitting}
\hl{To enable shifts of the solar wind frequency distributions with the SSN } , we add a linear SSN dependency to the median
\begin{align}
	x_\text{med}(ssn) &= a_\text{med} \cdot ssn + b_\text{med}\,,	\label{eq:median_with_ssn}
\end{align}
using a factor to the SSN $a_\text{med}$ with a baseline $b_\text{med}$. We relate the mean with a scaling factor to the median to transfer its SSN dependency:
\begin{align}
	x_\text{avg}(ssn) &= (1 + a_\text{avg}) \cdot x_\text{med}(ssn)\,.	\label{eq:mean_with_ssn}
\end{align}
\hl{With the implementation of these relations into the lognormal function } (\ref{eq:single_lognormal_fit_function}), the new dynamic fit function $W'(x,ssn)$ is then fitted to the yearly data. The three resulting fit coefficients ($a_\text{med}, b_\text{med}$ and $a_\text{avg}$) are presented in Table~\ref{tab:ssn_fit_parameters}.
\begin{table*}
	\caption{Resulting fit coefficients from the OMNI data fitting with lagged SSN. For the velocity the fit parameters from the double lognormal fit and their balancing function are given. The values in brackets are the \hl{estimated standard deviation} of the fit parameters.}
	\label{tab:ssn_fit_parameters}
	\centering
	\begin{tabular}{l@{} c@{}
		S[table-format = 1.3(2)e+1]
		S[table-format = 1.4(2)]
		S[table-format = 1.3(2)e+1]
		S[table-format = +1.3(2)e+1]
		c
		}
		\hline\hline
		\multicolumn{2}{l}{\multirow{2}{*}{Parameter}}	&\multicolumn{2}{c}{Median\tablefootmark{a}}	&\multicolumn{1}{c}{Mean}	&\multicolumn{2}{c}{Balance}\\
		\cline{3-4}\cline{6-7}
		\multicolumn{2}{l}{}	&\multicolumn{1}{c}{SSN factor $a_\text{med}$}	&\multicolumn{1}{c}{Baseline $b_\text{med}$}	&\multicolumn{1}{c}{Scaling factor $a_\text{avg}$}	&\multicolumn{1}{c}{SSN factor $c_a$}	&\multicolumn{1}{c}{Baseline $c_b$}\\
		\hline
		\multicolumn{2}{l}{Magnetic field}	&1.309(19)e-2	&4.285(17)	&8.786(78)e-2	&\multicolumn{1}{c}{--}	&--\\
%		\multicolumn{2}{l}{Velocity}	&2.514(93)e-3	&3.8521(90)	&2.392(26)e-2	&\multicolumn{1}{c}{--}	&--\\
		\multicolumn{2}{l}{Density}	&3.81(25)e-3	&4.495(26)	&3.050(27)e-1	&\multicolumn{1}{c}{--}	&--\\
		\multicolumn{2}{l}{Temperature}	&1.974(26)e-2	&5.729(19)	&6.541(28)e-1	&\multicolumn{1}{c}{--}	&--\\
		\hline
		\multirow{2}{*}{Velocity}	&\multicolumn{1}{c}{$W'_1$}	&\multicolumn{1}{c}{--}	&3.633(12)	&1.008(37)e-2	&\multirow{2}{*}{\tablenum{-1.799(95)e-3}}	&\multirow{2}{*}{0.638(32)}\\
			&\multicolumn{1}{c}{$W'_2$}	&\multicolumn{1}{c}{--}	&4.831(81)	&2.31(20)e-2	&	&\\
		\hline
	\end{tabular}
	\tablefoot{
		\tablefoottext{a}{\hl{In units of } \si{nT}, \SI{e2}{\km\per\s}, \si{\per\cm\cubed} and \SI{e4}{\K}.}
	}
\end{table*}

@@

Again, the solar wind velocity needs a special treatment because of the application of the double lognormal distribution (\ref{eq:double_lognormal_fit_function}). Since it is well known that slow and fast solar wind stream occurrence rates follow the solar cycle we keep the two velocity components' positions constant and vary instead their balance with the SSN:
\begin{align}
	c(ssn) &= c_a \cdot ssn + c_b\,.	\label{eq:balance_with_ssn}
\end{align}
The fit result (see Table~\ref{tab:ssn_fit_parameters}) yields a model in which three years after solar cycle minimum (SSN of zero) the contribution of slow solar wind to the overall solar wind distribution reaches a maximum value of about \SI{64}{\percent} and decreases with increasing SSN as shown in Fig.~\ref{fig:Vdbl_SSN_ratio_f_plot}.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/Vdbl_SSN_ratio_f_plot.pdf}}
	\caption{Ratio of slow to fast solar wind for a by 3~years lagged SSN. The yearly ratios (+) and their weighted linear fit (solid line) are obtained from OMNI data with a threshold velocity of $v_\text{th} = \SI{400}{\km\per\s}$. The error bars denote the SSN standard deviation and the relative weight from the yearly data coverage. The model's balance parameter (\ref{eq:balance_with_ssn}) and derived ratio (same threshold) are plotted as dashed and dotted lines.}
	\label{fig:Vdbl_SSN_ratio_f_plot}
\end{figure}
% 3 years after minimum (SSN of zero) the slow wind has its maximal share of 0.638(32). 3 years after maximum (SSN of 200) it has only a share of 0.278(37).

@@

The median and mean values of the key solar wind parameters for different solar distances of the Helios data are calculated for the minimal distance resolution \SI{0.01}{\au} of the data set\hl{, see } Fig.~\ref{fig:radial_fit_4_thesis_light_b_skip}. Assuming a radial solar wind outflow, it is expected that the distance dependence of the solar wind parameters over the Helios data range \SIrange{0.29}{0.98}{\au} can be described through power law scaling. Therefore we use the power law function
\begin{align}
	x(r) = d\cdot r^e	\label{eq:power_function}
\end{align}
for the regression fit of the median and mean, with $r$ being the solar distance in astronomical units, $d$ the magnitude at \SI{1}{\au} and $e$ the exponent. The fits are weighted through the different data counts per bin.
%fit result table and figure
The fit coefficients ($d_\text{med}$, $d_\text{avg}$, $e_\text{med}$ and $e_\text{avg}$) are listed in Table~\ref{tab:mean_median_fit_parameter}.
\begin{table*}
	\caption{Fit coefficients for \hl{the median and mean solar distance dependencies } of the four solar wind parameters derived from the combined Helios~1 and 2 data. The errors in brackets are the \hl{estimated standard deviations} of each fit parameter. The crossing distances indicate where the median and mean fits intersect each other. The yearly variation is the weighted standard deviation derived from the yearly fit exponents seen in Fig.~\ref{fig:yearly_gradients_c}.}
	\label{tab:mean_median_fit_parameter}
	\centering
	\begin{tabular}{l
	S[table-format = 1.3(2)]
	S[table-format = +1.3(2)]
	@{}c
	S[table-format = 1.3(2)]
	S[table-format = +1.3(2)]
	S[table-format = 1.1(2)e1]
	S[table-format = 1.3]}
		\hline\hline
		\multirow{2}{*}{Parameter}	&\multicolumn{2}{c}{Median}	&	&\multicolumn{2}{c}{Mean}	&\multicolumn{1}{c}{Crossing distance}	&\multicolumn{1}{c}{Yearly variation}\\
		\cline{2-3}	\cline{5-6}
			&\multicolumn{1}{c}{$d_\text{med}$\tablefootmark{a}}	&\multicolumn{1}{c}{$e_\text{med}$}	&	&\multicolumn{1}{c}{$d_\text{avg}$\tablefootmark{a}}	&\multicolumn{1}{c}{$e_\text{avg}$}	&\multicolumn{1}{c}{[\si{\au}]}	&\multicolumn{1}{c}{$\Delta e$}\\
		\hline
		Magnetic field	&5.377(92)	&-1.655(17)	&	&6.05(10)	&-1.546(18)	&0.339(11)	&0.11\\
		Velocity	&4.107(28)	&0.058(13)	&	&4.356(24)	&0.049(10)	&0.7(83)e3	&0.012\\
		Density		&5.61(27)	&-2.093(46)	&	&7.57(30)	&-2.010(38)	&0.027(73)	&0.072\\
		Temperature	&7.14(23)	&-0.913(39)	&	&9.67(21)	&-0.792(28)	&0.082(85)	&0.050\\
		\hline
	\end{tabular}
	\tablefoot{
		\tablefoottext{a}{\hl{In units of } \si{nT}, \SI{e2}{\km\per\s}, \si{\per\cm\cubed} and \SI{e4}{\K}.}
	}
\end{table*}
% source year variation: fit_out_yearly_gradients_table_b_stats.txt

@@

%\subsection{Distance scaling law variations}
%\label{sec:distance_scaling_law_variations}
\textbf{This solar wind model} for the radial distance dependence is representative for the time of the Helios observations around the rise of solar cycle~21. The variation of yearly \hl{power law fit} exponents are shown in Fig.~\ref{fig:yearly_gradients_c} together with the yearly SSN for the time period \numrange{1974}{1982}. It can be seen that during the Helios time period there might be some systematic  variation of the exponents with solar activity---at least for the velocity and temperature exponents.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/yearly_gradients_c.pdf}}
	\caption{Helios yearly variation of the solar wind \textbf{parameter power exponents for the dependence on radial distance} together with the SIDC 13-month smoothed monthly SSN. The weighted standard deviations and average values for all years are indicated by the shaded areas. In this plot the 21~days since Helios launch in the year 1974 are omitted because a distance range of merely \SIrange{0.95}{0.98}{\au} was covered that year.}
	\label{fig:yearly_gradients_c}
\end{figure}
%yearly variation of the solar wind parameters' average fit function exponents
However, for simplicity we assume, that the distance scaling laws can be treated as time independent and include the calculated exponents’ yearly variations $\Delta e$, summarized in Table~\ref{tab:mean_median_fit_parameter}, as relative uncertainties.

@@

This model averages over seasonal variations because it is based on yearly data. As the OMNI data are time-shifted to the nose of the Earth’s bow shock, this leads to yearly solar distance variations of \SI{+-1.67}{\%} as it orbits the Sun. The resulting maximal solar wind parameter variation amplitudes over the year can thus be derived from the derived power law exponents. They are estimated to be smaller than \SI{4}{\percent} as seen in Tab.~\ref{tab:extrapolation_model_fit_parameters}. \citet{Bruno1986,Balogh1999} have pointed out, that the solar wind parameters vary with latitudinal separation from the heliospheric current sheet (HCS). Its position in heliographic latitude is highly variable around the solar equator \citep{Schwenn1990}, furthermore, the Earth’s orbit varies over the course of the year by \SI{+-7.2}{\degree} in latitude. \hl{However, the additional analysis of this aspect is beyond the scope of this study. } 	%Balogh et al. (p.~162, 1999);	 (Schwenn 1990, p.~126)

@@

%\subsection{Near-Sun extrapolation for PSP orbit}
%PSP orbit
Parker Solar Probe is planned to launch in mid 2018. With its first Venus flyby it will swing into Venus' orbital plane, reaching already 93~days after launch in November 2018 a first perihelion with a distance of \SI{0.16}{\au}. Seven additional Venus flybys allow \textbf{the perihelion distance to be reduced} to a minimum of \SI{9.86}{\Rs} \citep{Fox2015} as plotted in \hl{Fig. } ~\ref{fig:SPP_orbit_predicted_SSN_overview_e_plot}.

@@

This difference is \hl{arguably } due to the previously mentioned (Sect.~\ref{sec:solar_distance_dependency}) \textbf{changing shape of the frequency distribution with heliocentric distance} , \hl{which for smaller distances deviates more from the } model’s lognormal distribution.

>>>>

\section{Introduction}
%solar wind beginnings
From observations of cometary tail fluctuations \citet{Biermann1951} inferred the presence of a continuous flow of particles from the Sun. With his theoretical solar wind model \citet{Parker1958} formulated the existence of the solar wind even before the first satellites measured it in situ in 1959 \citep{Gringauz1960,Neugebauer1966}.
%in situ spacecraft
The idea of a space mission flying through the solar corona dates back to the founding year of NASA in 1958 \citep{McComas2008}. Since then several space missions have measured the solar wind in situ at a wide range of heliocentric distances, in case of Voyager~1 as far away as \SI{138}{\au}\footnote{\url{https://voyager.jpl.nasa.gov/}} (July 2017), having even left the heliospause into interstellar space at a distance of \SI{121}{\au} \citep{Gurnett2013}.
\textbf{Various} spacecraft have provided a wealth of solar wind measurements near Earth’s orbit, with WIND \citep{Lepping1995,Ogilvie1995}\textbf{, \hl{SOHO} \citep{Domingo1995}}, ACE \citep{Stone1998} and DSCOVR \citep{Burt2012} still orbiting around the L1 point \SI{1.5}{million\,\km} ahead of Earth in the sunward direction. Additional measurements at other solar distances were provided by planetary missions to Venus and Mercury, such as PVO \citep{Colin1980} or MESSENGER \citep{Belcher1991}. Ulysses was the first probe that orbited the Sun out of the ecliptic plane and thus could measure solar wind even at polar latitudes \citep{McComas1998}. \textbf{The in situ solar wind measurements closest to the Sun to date were made by} the Helios mission. \textbf{Helios~1, launched in 1974, reached distances of \SI{0.31}{\au}. Helios~2, launched two years later, approached the Sun as close as} \SI{0.29}{\au} \citep{Rosenbauer1977}.
%Solar Probe Plus mission
The NASA Parker~Solar~Probe\footnote{\url{http://parkersolarprobe.jhuapl.edu/}} (PSP), formerly Solar~Probe~Plus, with a planned launch date in mid 2018, will reach after six years in 2024 its closest perihelia at a distance of 9.86~solar radii (\Rs), that is, \SI{0.0459}{\au} \citep{Fox2015}. This distance will be achieved through seven Venus gravity assists with orbital periods of 88--168~days. In its prime mission time 2018--2025 PSP provides 24~orbits with perihelia inside \SI{0.25}{\au} \citep{Fox2015}. Even its first perihelion, 93~days after launch in 2018, will take PSP to an unprecedented distance of \SI{0.16}{\au} (\SI{35.7}{\Rs}). In comparison, the ESA Solar Orbiter mission with a planned launch in February 2019 will have its closest perihelia at \SI{0.28}{\au} \citep{Muller2013}.

@@

Fig.~\ref{fig:OMNI_yearly_ssn_correlation_c_plot} shows yearly medians of the solar wind parameters and the yearly SSN together with the solar cycle number. The reason for correlating the SSN to the solar wind median values is because the position of a lognormal function is defined by its median. The data are averaged to yearly values to avoid seasonal effects during the Earth’s orbit around the Sun caused by its variations in solar latitude and distance.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/OMNI_yearly_ssn_correlation_c_plot.pdf}}
	\caption{The solar wind parameter yearly \textbf{frequencies (grey shading) with yearly medians (lines)} derived from OMNI data and the yearly SSN from the \citet{sidc} with solar cycle number (top). Their correlation coefficients with the yearly SSN are calculated for time lags back to -15 years (bottom).}
	\label{fig:OMNI_yearly_ssn_correlation_c_plot}
\end{figure}
 Solar wind parameter yearly data frequencies
The solar wind velocity, density and temperature depend on the state of the solar cycle \citep{Schwenn1983}. %density modulation with SSN \citep{Schwenn1983} p.~499\\
For instance the fast solar wind \textbf{occurs at times when polar coronal holes extend to lower latitudes, a typical feature of the declining phase of the solar cycle} as pointed out by \citet[p.~75, Figure~3.52]{Bothmer2007}. Therefore the solar wind velocity, density and temperature maxima exhibit \textbf{time lags relative} to the SSN maxima.

The correlation coefficients of the solar wind parameters with the yearly SSN shown in the bottom part of Fig.~\ref{fig:OMNI_yearly_ssn_correlation_c_plot} are calculated for time lags back to \num{-15}~years to cover a time span longer than a solar cycle. As expected, the \textbf{amplitudes of the variations in the correlations} of all parameters decline with increasing time lag and show a \textbf{period} of about 11~years. The highest correlation coefficient of 0.728 to the SSN is found for the magnetic field strength, it has no time lag. This finding is anticipated because the SSN is found to be directly proportional to the evolution of the photospheric magnetic flux \citep{Smith2003}.
Velocity and temperature show a \textbf{time lag} of 3~years with peak correlation coefficients of 0.453 and 0.540. The density with a correlation coefficient of 0.468 has a time lag of 6~years, which is in agreement with the by \citet{Bougeret1984} reported density anticorrelation to the SSN.	%p.~406\\
% maximal correlation coefficients and their time lags:\\
% magnetic field strength: lag 0, 0.728\\
% velocity: lag -3, 0.453\\
% density: lag -6, 0.468\\
% temperature: lag -3, 0.540\\

%\subsection{SSN fitting}
\hl{To enable shifts of the solar wind frequency distributions with the SSN } , we add a linear SSN dependency to the median
\begin{align}
	x_\text{med}(ssn) &= a_\text{med} \cdot ssn + b_\text{med}\,,	\label{eq:median_with_ssn}
\end{align}
using a factor to the SSN $a_\text{med}$ with a baseline $b_\text{med}$. We relate the mean with a scaling factor to the median to transfer its SSN dependency:
\begin{align}
	x_\text{avg}(ssn) &= (1 + a_\text{avg}) \cdot x_\text{med}(ssn)\,.	\label{eq:mean_with_ssn}
\end{align}
\hl{With the implementation of these relations into the lognormal function } (\ref{eq:single_lognormal_fit_function}), the new dynamic fit function $W'(x,ssn)$ is then fitted to the yearly data. The three resulting fit coefficients ($a_\text{med}, b_\text{med}$ and $a_\text{avg}$) are presented in Table~\ref{tab:ssn_fit_parameters}.
\begin{table*}
	\caption{Resulting fit coefficients from the OMNI data fitting with \textbf{the linear SSN dependencies (\ref{eq:median_with_ssn}) and (\ref{eq:mean_with_ssn})}. For the velocity the fit parameters from the double lognormal fit \textbf{(\ref{eq:double_lognormal_fit_function})} and their balancing function \textbf{(\ref{eq:balance_with_ssn})} are given. The values in brackets are the \hl{estimated standard deviation} of the fit parameters.}
	\label{tab:ssn_fit_parameters}
	\centering
	\begin{tabular}{l@{} c@{}
		S[table-format = 1.3(2)e+1]
		S[table-format = 1.4(2)]
		S[table-format = 1.3(2)e+1]
		S[table-format = +1.3(2)e+1]
		c
		}
		\hline\hline
		\multicolumn{2}{l}{\multirow{2}{*}{Parameter}}	&\multicolumn{2}{c}{Median\tablefootmark{a}}	&\multicolumn{1}{c}{Mean}	&\multicolumn{2}{c}{Balance}\\
		\cline{3-4}\cline{6-7}
		\multicolumn{2}{l}{}	&\multicolumn{1}{c}{SSN factor $a_\text{med}$}	&\multicolumn{1}{c}{Baseline $b_\text{med}$}	&\multicolumn{1}{c}{Scaling factor $a_\text{avg}$}	&\multicolumn{1}{c}{SSN factor $c_a$}	&\multicolumn{1}{c}{Baseline $c_b$}\\
		\hline
		\multicolumn{2}{l}{Magnetic field}	&1.309(19)e-2	&4.285(17)	&8.786(78)e-2	&\multicolumn{1}{c}{--}	&--\\
%		\multicolumn{2}{l}{Velocity}	&2.514(93)e-3	&3.8521(90)	&2.392(26)e-2	&\multicolumn{1}{c}{--}	&--\\
		\multicolumn{2}{l}{Density}	&3.81(25)e-3	&4.495(26)	&3.050(27)e-1	&\multicolumn{1}{c}{--}	&--\\
		\multicolumn{2}{l}{Temperature}	&1.974(26)e-2	&5.729(19)	&6.541(28)e-1	&\multicolumn{1}{c}{--}	&--\\
		\hline
		\multirow{2}{*}{Velocity}	&\multicolumn{1}{c}{$W'_1$}	&\multicolumn{1}{c}{--}	&3.633(12)	&1.008(37)e-2	&\multirow{2}{*}{\tablenum{-1.799(95)e-3}}	&\multirow{2}{*}{0.638(32)}\\
			&\multicolumn{1}{c}{$W'_2$}	&\multicolumn{1}{c}{--}	&4.831(81)	&2.31(20)e-2	&	&\\
		\hline
	\end{tabular}
	\tablefoot{
		\tablefoottext{a}{\hl{In units of } \si{nT}, \SI{e2}{\km\per\s}, \si{\per\cm\cubed} and \SI{e4}{\K}.}
	}
\end{table*}

@@

Again, the solar wind velocity needs a special treatment because of the application of the double lognormal distribution (\ref{eq:double_lognormal_fit_function}). Since it is well known that slow and fast solar wind stream occurrence rates follow the solar cycle we keep the two velocity components' positions \textbf{SSN-independent ($x_\text{med} =  b_\text{med}$)} and vary instead their balance with the SSN:
\begin{align}
	c(ssn) &= c_a \cdot ssn + c_b\,.	\label{eq:balance_with_ssn}
\end{align}
The fit result (see Table~\ref{tab:ssn_fit_parameters}) yields a model in which three years after solar cycle minimum (SSN of zero) the contribution of slow solar wind to the overall solar wind distribution reaches a maximum value of about \SI{64}{\percent} and decreases with increasing SSN as shown in Fig.~\ref{fig:Vdbl_SSN_ratio_f_plot}.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/Vdbl_SSN_ratio_f_plot.pdf}}
	\caption{Ratio of slow to fast solar wind for a by 3~years lagged SSN. The yearly ratios (+) and their weighted linear fit (solid line) are obtained from OMNI data with a threshold velocity of $v_\text{th} = \SI{400}{\km\per\s}$. The error bars denote the SSN standard deviation and the relative weight from the yearly data coverage. The model's balance parameter (\ref{eq:balance_with_ssn}) and derived ratio (same threshold) are plotted as dashed and dotted lines.}
	\label{fig:Vdbl_SSN_ratio_f_plot}
\end{figure}
% 3 years after minimum (SSN of zero) the slow wind has its maximal share of 0.638(32). 3 years after maximum (SSN of 200) it has only a share of 0.278(37).

@@

The median and mean values of the key solar wind parameters for different solar distances of the Helios data are calculated for the minimal distance resolution \SI{0.01}{\au} of the data set\hl{, see } Fig.~\ref{fig:radial_fit_4_thesis_light_b_skip}. Assuming a radial solar wind outflow, it is expected that the distance dependence of the solar wind parameters over the Helios data range \SIrange{0.29}{0.98}{\au} can be described through power law scaling. Therefore we use the power law function
\begin{align}
	x(r) = d\cdot r^e	\label{eq:power_function}
\end{align}
for the regression fit of the median and mean, with $r$ being the solar distance in astronomical units, $d$ the magnitude at \SI{1}{\au} and $e$ the exponent. The fits are weighted through the different data counts per bin.
%fit result table and figure
The fit coefficients ($d_\text{med}$, $d_\text{avg}$, $e_\text{med}$ and $e_\text{avg}$) are listed in Table~\ref{tab:mean_median_fit_parameter}.
\begin{table*}
	\caption{Fit coefficients for the median and mean solar distance dependencies \textbf{(\ref{eq:power_function})} of the four solar wind parameters derived from the combined Helios~1 and 2 data. The errors in brackets are the \hl{estimated standard deviations} of each fit parameter. The crossing distances indicate where the median and mean fits intersect each other. The yearly variation is the weighted standard deviation derived from the yearly fit exponents seen in Fig.~\ref{fig:yearly_gradients_c}.}
	\label{tab:mean_median_fit_parameter}
	\centering
	\begin{tabular}{l
	S[table-format = 1.3(2)]
	S[table-format = +1.3(2)]
	@{}c
	S[table-format = 1.3(2)]
	S[table-format = +1.3(2)]
	S[table-format = 1.1(2)e1]
	S[table-format = 1.3]}
		\hline\hline
		\multirow{2}{*}{Parameter}	&\multicolumn{2}{c}{Median}	&	&\multicolumn{2}{c}{Mean}	&\multicolumn{1}{c}{Crossing distance}	&\multicolumn{1}{c}{Yearly variation}\\
		\cline{2-3}	\cline{5-6}
			&\multicolumn{1}{c}{$d_\text{med}$\tablefootmark{a}}	&\multicolumn{1}{c}{$e_\text{med}$}	&	&\multicolumn{1}{c}{$d_\text{avg}$\tablefootmark{a}}	&\multicolumn{1}{c}{$e_\text{avg}$}	&\multicolumn{1}{c}{[\si{\au}]}	&\multicolumn{1}{c}{$\Delta e$}\\
		\hline
		Magnetic field	&5.377(92)	&-1.655(17)	&	&6.05(10)	&-1.546(18)	&0.339(11)	&0.11\\
		Velocity	&4.107(28)	&0.058(13)	&	&4.356(24)	&0.049(10)	&0.7(83)e3	&0.012\\
		Density		&5.61(27)	&-2.093(46)	&	&7.57(30)	&-2.010(38)	&0.027(73)	&0.072\\
		Temperature	&7.14(23)	&-0.913(39)	&	&9.67(21)	&-0.792(28)	&0.082(85)	&0.050\\
		\hline
	\end{tabular}
	\tablefoot{
		\tablefoottext{a}{\hl{In units of } \si{nT}, \SI{e2}{\km\per\s}, \si{\per\cm\cubed} and \SI{e4}{\K}.}
	}
\end{table*}
% source year variation: fit_out_yearly_gradients_table_b_stats.txt

@@

%\subsection{Distance scaling law variations}
%\label{sec:distance_scaling_law_variations}
\textbf{This solar wind model} for the radial distance dependence is representative for the time of the Helios observations around the rise of solar cycle~21. The variation of yearly power law fit exponents \textbf{by fitting the solar distance dependency (\ref{eq:power_function})} are shown in Fig.~\ref{fig:yearly_gradients_c} together with the yearly SSN for the time period \numrange{1974}{1982}. It can be seen that during the Helios time period there might be some systematic  variation of the exponents with solar activity---at least for the velocity and temperature exponents.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/yearly_gradients_c.pdf}}
	\caption{Helios yearly variation of the solar wind \textbf{parameter power exponents for the dependence on radial distance} together with the SIDC 13-month smoothed monthly SSN. The weighted standard deviations and average values for all years are indicated by the shaded areas. In this plot the 21~days since Helios launch in the year 1974 are omitted because a distance range of merely \SIrange{0.95}{0.98}{\au} was covered that year.}
	\label{fig:yearly_gradients_c}
\end{figure}
%yearly variation of the solar wind parameters' average fit function exponents
However, for simplicity we assume, that the distance scaling laws can be treated as time independent and include the calculated exponents’ yearly variations $\Delta e$, summarized in Table~\ref{tab:mean_median_fit_parameter}, as relative uncertainties.

@@

This model averages over seasonal variations because it is based on yearly data. As the OMNI data are time-shifted to the nose of the Earth’s bow shock, this leads to yearly solar distance variations of \SI{+-1.67}{\%} as it orbits the Sun. The resulting maximal solar wind parameter variation amplitudes over the year can thus be derived from the derived power law exponents. They are estimated to be smaller than \SI{4}{\percent} as seen in Tab.~\ref{tab:extrapolation_model_fit_parameters}. \citet{Bruno1986} \textbf{and} \citet{Balogh1999} have pointed out, that the solar wind parameters vary with latitudinal separation from the heliospheric current sheet (HCS). Its position in heliographic latitude is highly variable around the solar equator \citep{Schwenn1990}, furthermore, the Earth’s orbit varies over the course of the year by \SI{+-7.2}{\degree} in latitude. \textbf{Since this latitudinal separation is highly variable, and requires significant effort to calculate for an extended time series, we have ignored this aspect in this analysis.}	%Balogh et al. (p.~162, 1999);	 (Schwenn 1990, p.~126)

@@

%\subsection{Near-Sun extrapolation for PSP orbit}
%PSP orbit
Parker Solar Probe is planned to launch in mid 2018. With its first Venus flyby it will swing into Venus' orbital plane, reaching already 93~days after launch in November 2018 a first perihelion with a distance of \SI{0.16}{\au}. Seven additional Venus flybys allow \textbf{the perihelion distance to be reduced} to a minimum of \SI{9.86}{\Rs} \citep{Fox2015} as plotted in \hl{Fig. } ~\ref{fig:SPP_orbit_predicted_SSN_overview_f_plot}.

@@

\textbf{We suggest that this difference is} due to the previously mentioned (Sect.~\ref{sec:solar_distance_dependency}) \textbf{changing shape of the frequency distribution with heliocentric distance} , \hl{which for smaller distances deviates more from the } model’s lognormal distribution.
